Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-14 ss-lin-EPQ.tex
change explanation

>>>>

Let us briefly explain the idea of proving step-by-step linearizability of $\textit{EPQ}_1$ with an example, while the other two rules is much simpler to deal with. Given a data-differentiated concurrent execution $e \sqsubseteq l = u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w \in \textit{MS}(\textit{EPQ}_1)$ with witness $x$ and assume that $e \setminus x \sqsubseteq l' \in \textit{EPQ}$, we explicitly construct a sequence $l''= l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$ and prove that $e \sqsubseteq l'' \in \textit{EPQ}$. $e$ is shown in \figurename~\ref{fig:concurrent execution for EPQ1} and we explicitly draw the linearization points according to $l'$. Here $\textit{pri}_x=p_2$, and assume that $p_1 <_{\mathbb{P}} p_2$ and $p_1 <_{\mathbb{P}} p_3$. Here $u = \epsilon$, $v = \textit{put}(z_1,p_3) \cdot \textit{put}(x_2,p_2) \cdot \textit{rm}(y_1) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(z_2) \cdot \textit{rm}(x_2)$, and $w=\textit{rm}(z_2) \cdot \textit{put}(x_3,p_2) \cdot \textit{rm}(z_1)$. We use $o-w$ to emphasize that $o$ is in $w$.

The construction of $l''$ is as follows. Our construction does not rely how to choose linearization points of $l'$ in $e$.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] At first glance, we can construct $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ into operations of $u$, $v$ and $w$, respectively. However, this is incorrect. To explain this, let $\textit{pri}_x$-comparable operations (resp., $\textit{pri}_x$-incomparable operations) be the operations with items whose priority is comparable with $\textit{pri}_x$ (resp., incomparable with $\textit{pri}_x$). According to $\textit{EPQ}_1$, there is no restriction to $\textit{pri}_x$-incomparable operations operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ into $\textit{pri}_x$-incomparable operations operations in $u \cdot v$ being correct. In this example, we can see that such projection is $\textit{put}(z_1,p_3) \cdot \textit{rm}(z_2)$ and is incorrect.

\item[-] Let us construct set $U'$, $V'$ and $W'$, such that $l''_1$, $l''_2$ and $l''_3$ are projection of $l'$ into $U'$, $V'$ and $W'$, respectively. The construction contains two steps:

\item[-] The first step is to define $W'$. The $\textit{pri}_x$-comparable operations in $W'$ is same as that in $W$. To obtain $\textit{pri}_x$-incomparable operations in $W'$, we try to find an operation $o$ which either happens before some $\textit{pri}_x$-comparable operations in $W$, or with linearization points after $\textit{rm}(x)$. In this example, $o$ is $\textit{put}(x_3,p_2)$ (emphasized by adding vertical dashed line). Then, we put $o$ and all $\textit{pri}_x$-incomparable operations in $l'$ whose linearization points are after $o$ into $W'$. In this example, $W'$ contains $\textit{put}(x_3,p_2)$, $\textit{rm}(z_1)$ and $\textit{rm}(z_2)$. We use boxes to emphasize they are put into $W'$.

\item[-] The second step is to define $U'$ and $V'$. $U'$ contains the following two kinds of operations: (1) Operations whose linearization points are before $\textit{ret}(\textit{put},x)$, and (2) other $\textit{put}$ operations with priority $\textit{pri}_x$. $V'$ contains the remanning operations. In this example, $U'$ contains $\textit{put}(z_1,p_3)$ and $\textit{put}(x_2,p_2)$.

\item[-] In this example, $l''_1 = \textit{put}(z_1,p_3) \cdot \textit{put}(x_2,p_2)$, $l''_2 = \textit{put}(z_2,p_3) \cdot \textit{rm}(x_2) \cdot \textit{put}(y_1,p_1) \cdot \textit{rm}(y_1)$, and $l''_3 = \textit{put}(x_3,p_2) \cdot \textit{rm}(z_1) \cdot \textit{rm}(z_2)$. In \figurename~\ref{fig:concurrent execution with new linearization points for EPQ1}, we add linarization points according to $l''$, and we can see that $l''$ holds as required.
\end{itemize}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5 \textwidth]{figures/PIC-HIS-EPQ1.pdf}
%\vspace{-10pt}
  \caption{Concurrent execution $e$}
  \label{fig:concurrent execution for EPQ1}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5 \textwidth]{figures/PIC-HIS-EPQ1-NEWLP.pdf}
%\vspace{-10pt}
  \caption{Concurrent execution $e$ with linearization points according to $l''$}
  \label{fig:concurrent execution with new linearization points for EPQ1}
\end{figure}
